---
output: 
  github_document:
    fig_width: 8
    fig_height: 10
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(tidyverse)
library(cowplot)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {cudpalette}

<!-- badges: start -->

[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)

<!-- badges: end -->

Create color palette function using model color palette for Color Universal Design (Color Universal Design Organization, 2018, https://cudo.jp/?page_id=1565) or JIS Z 9103:2018.

## About this package (in Japanese) 

CUDO（NPO法人 カラーユニバーサルデザイン機構）により示されている「カラーユニバーサルデザイン推奨配色セット ver.4」(https://cudo.jp/?page_id=1565) および「カラーユニバーサルデザイン推奨配色セット ガイドブック 第2版」（[PDF](https://cudo.jp/wp-content/uploads/2018/10/cud_guidebook.pdf)）の「比較的見分けやすい組み合わせ（画面用）」をカラーパレットとして利用できるようにする。

同様の趣旨で制定されている[JIS安全色](https://safetycolor.jp/)（JIS Z 9103:2018）も使用できるよう更新。

## Functions :

### `cud_color(colname)`

Get Color Universal Design or JIS Z 9103:2018 color code(s) from color name(s) listed berow. The color name(s) for JIS color should be specified with a "jis_" prefix. \
If "colorname" is not exist in the list, `NA` will be return.

> CUD推奨カラーパレットの色名（下記）から対応するカラーコードを取得。存在しない色は NA を返す

```{r, echo = FALSE}
names(cudpalette:::list_cud_cols)
```

> `jis_` の接頭語をつけて `jis_color()` の色を取得することもできる

```{r, echo = FALSE}
paste0("jis_", names(cudpalette:::list_jis_cols))
```


### `jis_color(colname)`

Get JIS Z 9103:2018 color code.\
`NA` will be return when "colorname" is not exist in the following list.

> JIS Z 9103:2018 で定義されているカラーコードを取得。存在しない色は NA を返す

```{r, echo = FALSE}
names(cudpalette:::list_jis_cols)
```


### `cud_pal(type, size, pal, direction, reorder, warn)`

Use patterns recommended as easy-to-distinguish combinations in the materials as a palette. See the sample at the end of this page for possible combinations.\
The default pallete is `type = "accent", size = 6, pal = 1`. If requested number of colors is more than in the pallete, colors are used repeatedly (warning message could be suppressed with `warn = F` option).\
Colors are reorderd to maximize the color difference between adjacent colors. If you want to go back to the order in which they described in the source material, use `reorder = F`.

> 「比較的見分けやすい組み合わせ」をパレットとして使用する。可能な組み合わせは下記サンプルを参照。\
隣り合う色の色差を最大化するために色を並べ替えている。出典資料に出てくる順に戻したい場合は、`reorder = F`。
>
> デフォルトはアクセントカラー 6色のひとつ。色が足りない場合は警告メッセージ（`warn = F` でoff可）を出してループで使用する。

### `jis_pal(direction, reorder, warn)`

This is a wrapper function calls 'cud_pal()' with type = "jis".

> JIS Z 9103:2018 で定義されている色をパレットとして使用。`cud_pal(type = "jis")` を呼び出すラッパー関数。

## Installation

``` r
remotes::install_github("mokztk/cudpalette")
```
## Example

### basic usage

```{r example}
library(cudpalette)

cud_color("red")

cud_color(c("red", "blue", "magenta", "purple"))

cud_pal()()

cud_pal(type = "a", size = 5, pal = 2, direction = -1)()

cud_pal(type = "a", size = 7, pal = 2, direction = -1)()

cud_pal(type = "a", size = 5, pal = 2, direction = -1)(6)

cud_pal(type = "a", size = 5, pal = 2, direction = -1, warn = FALSE)(6)
```

### color sample

with parameter `reorder = TRUE` (reorder palettes maximizing color differences between adjacent colors)

```{r sample_graph, echo = FALSE}
library(cudpalette)
library(tidyverse)
library(cowplot)

tbl_pallete <-
  tribble(
    ~type,         ~size, ~pal_max,
    "accent",      6,     3,
    "accent",      5,     5,
    "accent",      4,     6,
    "base",        4,     2,
    "base",        3,     5,
    "accent_base", 6,     8,
    "jis",         6,     1,
  ) %>%
  mutate(pal = map(pal_max, ~ 1:.x)) %>%
  select(-pal_max) %>%
  unnest(cols = pal) %>% 
  mutate(
    palette = pmap(list(type, size, pal),
                   function(type, size, pal) cud_pal(type, size, pal, warn = F)())
  )

set.seed(23456)
sample_graph <-
  tibble(
    x = LETTERS[1:6],
    y = runif(6, 6, 10)
  ) %>% 
  ggplot(aes(x, y, fill = x, linetype = x)) +
    geom_col(colour = "black") +
    scale_y_continuous(limits = c(0, 10), expand = c(0, 0)) +
    guides(fill = "none", linetype = "none") + 
    theme_light() + 
    theme(
      axis.line = element_line(colour = "black"),
      axis.text = element_blank(), 
      axis.title = element_blank(),
      plot.title = element_text(size = 9)
    )

tbl_pallete %>% 
  mutate(
    plot = pmap(list(type, size, pal, palette),
                function(type, size, pal, palette) 
                  sample_graph + 
                    ggtitle(paste(type, "/ size =", size, "/ pal =", pal)) +
                    scale_fill_manual(values = c(palette, rep("white", 6))) +
                    scale_linetype_manual(values = c(rep("solid", length(palette)),
                                                     rep("dotted", 6)))
                )
  ) %>%
  pull(plot) %>% 
  plot_grid(plotlist = ., ncol = 4)
```

-----

# Version history

- :bookmark:[v0.1.0](https://github.com/mokztk/cudpalette/releases/tag/v0.1.0) (2023-10-12): impremented the model color palette for Color Universal Design.
- :bookmark:[v0.2.0](https://github.com/mokztk/cudpalette/releases/tag/v0.2.0) (2023-10-20): impremented the JIS Safety Colors (JIS Z 9103:2018).
- :bookmark:[v0.3.0](https://github.com/mokztk/cudpalette/releases/tag/v0.3.0) (2023-10-20): reorder palettes to maximize color differences between adjacent colors.
- :bookmark:[v0.4.0](https://github.com/mokztk/cudpalette/releases/tag/v0.4.0) (2024-02-20): reorder (again) palettes to maximize color differences considering repeated use.
